### 2016-01-31 paper/ms.tex
added manuscript, obs: version after first ref report

%                                                                 aa.dem
% AA vers. 8.2, LaTeX class for Astronomy & Astrophysics
% demonstration file
%                                                       (c) EDP Sciences
%-----------------------------------------------------------------------
%
%\documentclass[referee]{aa} % for a referee version
%\documentclass[onecolumn]{aa} % for a paper on 1 column  
%\documentclass[longauth]{aa} % for the long lists of affiliations 
%\documentclass[rnote]{aa} % for the research notes
%\documentclass[letter]{aa} % for the letters 
%\documentclass[bibyear]{aa} % if the references are not structured 
% according to the author-year natbib style
%
\documentclass{aa}

%
\usepackage{graphicx}

\usepackage{hyperref}
\hypersetup{
    bookmarks=true,                                  % show bookmarks bar in pdf reader
    pdftitle={A Consistent 10 Myr Age for the Upper Scorpius OB Association},          % set pdf title          
    pdfauthor={Gregory A. Feiden},   % set pdf author
    colorlinks=true,                                 % false = box link, true = colored links
    linkcolor=blue,                                  % color of internal links
    citecolor=blue,                                  % color of citations
    urlcolor=blue                                    % external url color
}

\usepackage{defs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{txfonts}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\citestyle{aa}
\bibliographystyle{aa}

\titlerunning{A Consistent 10 Myr Age for Upper Scorpius}
\authorrunning{G.~A. Feiden}

\begin{document}

   \title{Magnetic Inhibition of Convection and the Fundamental Properties of Low--Mass Stars}
   \subtitle{III. A Consistent 10 Myr Age for the Upper Scorpius OB Association}

   \author{Gregory A.~Feiden}

   \institute{Department of Physics \& Astronomy, Uppsala University, Box 516,
                SE-751 20, Uppsala, Sweden. \\
                \email{gregory.feiden@physics.uu.se}
             }

   \date{Version: \today; Submitted: October 21, 2015; Re-submitted: February DD, 2016.}

  \abstract{
    When determining absolute ages of identifiably young stellar populations, results strongly depend on which stars are studied. Cooler (K, M) stars typically yield ages that are systematically younger than warmer (A, F, G) stars, typically by a factor of two. I explore the possibility that this age discordance is the product of magnetic inhibition of convection in cool young stars. I demonstrate that the adoption of magnetic stellar evolution models yield a median age of {\bf 9 -- 10} Myr for the Upper Scorpius subgroup of the Scorpius-Centaurus OB Association. This age is consistent for stars with spectral types A through M, except for a subset of F-type stars that appear older. Agreement is shown for ages derived from the Hertzsprung-Russell diagram and using an empirical mass-radius relationship from eclipsing multiple-star systems. Surface magnetic field strengths required to produce agreement are {\bf of order 2.5 kG} and are predicted from \emph{a priori} estimates of equipartition values, meaning no fine-tuning of model parameters is necessary. {\bf A gradual convergence of magnetic isochrones with standard model isochrones of an equivalent age is predicted to occur with the thinning of convective envelopes in stars with masses above 1.2~$M_{\odot}$. The predicted convergence at 10 Myr occurs when the mass contained in the convective envelope is approximately equal to 40\% of the total stellar mass. This transition appears to coincide with a noted flattening of the Hertzsprung-Russell diagram morphology of Upper Scorpius.} Furthermore, depletion of lithium predicted from magnetic models appears in better agreement with observed lithium equivalent widths than predictions from non-magnetic models. These results suggest that magnetic inhibition of convection plays an important role in the early evolution of low-mass stars and that it may be responsible for noted age discrepancies in young stellar populations.}

   \keywords{stars: evolution -- stars: low-mass -- stars: magnetic fields -- stars: pre-main-sequence -- open clusters and associations: individual: Upper Scorpius}

   \maketitle
%
%________________________________________________________________

\section{Introduction}
The \object{Upper Scorpius} subgroup of the larger Scorpius-Centaurus OB association is estimated to have a median age {\bf between 5 and 11 Myr.} Prior to 2012, there was a broad consensus that Upper Scorpius is roughly 5 Myr old based on age determinations from stellar evolution models \citep{deGeus1992, Preibisch2002, Slesnick2008}. Crucially, \citet{Preibisch2002} showed that this age was consistent across the entire theoretical Hertzsprung-Russell (HR) diagram. However, it was quickly suggested that early-type stars in Upper Scorpius supported an older 8 -- 10 Myr age \citep{Sartori2003}. This revision was confirmed by \citet{Pecaut2012} who performed a comprehensive re-analysis of the B-, A-, F-, and G-type members of Upper Scorpius and found HR diagram ages of each stellar spectral type are consistent with an overall age of approximately $11\pm3$ Myr for the subgroup. Despite the older median age inferred from high mass members of Upper Scorpius, ages for cooler K- and M-type members remain steadfastly younger at 5 -- 7 Myr \citep[][]{Herczeg2015, Rizzuto2015, Rizzuto2015b}, incompatible with a proposed age of 11 Myr.

This problem is not unique to Upper Scorpius and is characteristic of broader issues endemic to age dating young stellar populations with stellar evolution isochrones \citep[e.g.,][]{Naylor2009, Bell2012, Herczeg2015}. HR diagram analyses by \citet{Malo2014} and \citet{Herczeg2015} show that age estimates of mid- to late-M stars in nearby young stellar moving groups are typically younger by factor of two as compared to early-M and late-K stars. On the other hand, early- to late-K stars appear to have ages that are roughly consistent with each other. However, these ages are another factor of two younger than ages typically inferred from stars with spectral type G or earlier \citep{Hillenbrand2008}.

This picture is consistent with findings from studies of open cluster color-magnitude diagrams. Pre-main-sequence stars typically appear a factor of 2 -- 5 younger than main-sequence stars in the same population \citep{Naylor2009, Bell2012, Bell2013}. Age determinations appear to be a sensitive function of the stellar effective temperature range adopted in the analysis. The question is whether this is intrinsically real or an artifact of observational or modeling errors.

%This paper focuses on the latter following evidence that introduction of magnetic fields in stellar models can slow the contraction of young, low-mass stars \citep[e.g.,][]{MM10, Malo2014}.
%% Add paragraph of background with D'Antona+; Malo+; MM10/15;

While other young stellar populations exhibit an age gradient as a function of effective temperature, Upper Scorpius poses a particularly interesting hurdle. Previous studies of this problem are confined to the theoretical HR diagram and suffer from significant uncertainties in the accuracy of inferred stellar properties. However, observations of Upper Scorpius by \emph{Kepler}/K2 during Campaign 2 provide accurate photometric lightcurves of eclipsing multi-star systems \citep{Kraus2015, Alonso2015, David2015b, Lodieu2015}. This opens up new territory for assessing hypotheses as to the origin of the noted age discordance. Models can now be constrained in the HR diagram with added constraints from the eclipsing binary mass-radius relationship. The empirical mass-radius relationship is more reliable than the HR diagram, as masses and radii can be measured directly \citep[e.g.,][]{Andersen1991,Torres2010}. Compare this to conversions from spectral types and photometric colors to \teff s and luminosities, which are subject to significant uncertainties, especially for stars with gravities intermediate between dwarf and giant scales \citep[see, e.g.,][]{Pecaut2013}. Empirical mass-radius relationships should therefore provide stringent tests of stellar model accuracy.

\citet{Kraus2015} examined stellar model agreement using the properties of a low-mass eclipsing binary, UScoCTIO5, that has two roughly 0.3 \msun\ stars with precisely measured radii. What they found was that \object{UScoCTIO5} has an estimated age of about 5 Myr if inferred from the HR diagram, consistent with the age of other low-mass stars in Upper Scorpius \citep{Preibisch2002, Slesnick2008}. Surprisingly, the age inferred for UScoCTIO5 from the mass-radius relationship implies UScoCTIO5 is about 8 Myr old, more in line with \citet{Pecaut2012}. {\bf There is some disagreement about the precise values for the masses and radii of UScoCTIO5 \citep{David2015b}, but the mass-radius relationship nevertheless suggests an age of 7 -- 8 Myr}. Stellar ages are not only dependent on the effective temperature of stars used in the analysis, but on the specific properties adopted. It appears that stellar evolution models are unable to provide consistent ages on multiple levels.

Unfortunately, a single eclipsing binary system with two near-equal-mass components only gives a narrow view of model reliability. More than one system is needed to draw more definitive conclusions. A second system was supplied by the characterization of the triply eclipsing hierarchical triple \object{HD 144548} \citep{Alonso2015}. HD 144548 provides three additional points that are located in a separate region of the mass-radius and HR diagrams than UScoCTIO5 (two near $1.0$ \msun\ and one near $1.4$ \msun). Characterization of HD 144548 in combination with UScoCTIO5 allows for stricter tests of model consistency, assessing not just the ability of models to reproduce individual binary stars, but the more general slope of the mass-radius relationship concurrently with HR diagram fitting.

Here, I test the hypothesis that magnetic fields may inhibit convection in low-mass pre-main-sequence stars {\bf slowing their contraction along the Hayashi track. Slower contraction times for low-mass stars} would give them larger radii at a given age, pushing inferred ages to older values \citep[e.g.,][]{MM10, Malo2014}.  
In Sect.~\ref{sec:models}, I describe the model setup and method used to predict stellar surface magnetic field strengths \emph{a priori}. Section~\ref{sec:data} presents the data adopted for use in HR and mass-radius diagrams. Results about the performance of standard and magnetic stellar evolution models are given in Sect.~\ref{sec:results}, followed by  a discussion of potential uncertainties and weaknesses, as well as further evidence in support of the conclusions in Sect.~\ref{sec:disc}. Finally, the main conclusions of the paper are summarized in Section~\ref{sec:tellit}.

\section{Stellar Evolution Models}
\label{sec:models}
Models adopted in the analysis are magnetic Dartmouth stellar evolution models \citep{FC12b}. Basic physics used in these models are similar to those included in the original Dartmouth Stellar Evolution Program \citep[DSEP;][]{Dotter2008}. There are noted improvements that allow for a more accurate treatment of young star evolution \citep[see, e.g.,][]{Malo2014}. Key improvements are the explicit treatment of deuterium burning in the nuclear reaction network and the prescription of surface boundary conditions at a location in the stellar envelope where the optical depth $\tau_{5000} = 10$.

Specification of surface boundary conditions requires the determination of the gas pressure $\pgas$ and gas temperature $T_{\rm gas}$ at a given optical depth as a function of \logg, \teff, and [$m$/H]. This is accomplished by extracting relevant quantities from stellar model atmosphere structures that are computed with the same solar abundance distribution and (hopefully) opacity data as the interior structure models. Here, model atmosphere structures are those from a custom PHOENIX AMES-COND grid \citep{Hauschildt1999a, Hauschildt1999b, Dotter2008} that adopts the solar composition of \citet{GS98} and low-temperature opacities of \citet{Ferguson2005}.

Effects related to the interaction of magnetic fields and convection, as well as magnetic fields on the plasma equation of state, are included as described by \citet{FC12b,FC13}. {\bf Radial magnetic field strength profiles are prescribed using a ``Gaussian magnetic field strength profile'' of the form
\begin{equation}
	B(r) = B(R_{\rm src}) \exp\left[-\frac{1}{2}\left(\frac{R_{\rm src} - r}{\sigma_g}\right)^2\right]
\end{equation}
where $R_{\rm src}$ is the location of the inner convection zone boundary in fractional radii and $\sigma_g$ defines the width of the Gaussian \citep{FC13}. The adopted radial magnetic field strength profile has a negligible impact on the results provided the peak magnetic energy density remains small compared to the thermal energy density \citep{FC13,FC14,FC14b}.}
Peak interior magnetic field strengths are defined at $R_{\rm src} = 0.5\ R_{\star}$ for fully convective models and at the maximum of $R_{\rm src} = 0.5\ R_{\star}$ and $R_{\rm src} = R_{\rm bcz}$ (convection zone boundary) for partially convective stars. This latter condition allows magnetic models smoothly transition from a fully convective to partially convective configuration.

{\bf The width of the Gaussian controls the magnitude of the peak magnetic field strength at the inner convection zone boundary. It was defined by \citet{FC13} as
\begin{equation}
	\sigma_g = 0.2264 - 0.1776 \left(R_{\rm src}/R_{\star}\right).
\end{equation}
\citet{FC13} provided this expression for $\sigma_g$ to keep the Gaussian localized (sharply peaked) in stars with thin convective envelopes and to make a more distributed (broadly peaked) magnetic field in fully convective stars. For stars with very thin convective envelopes or in stars where the convection envelope gradually disappears, $\sigma_g$ becomes increasingly small and produces large peak magnetic field strengths at the convection zone boundary. As a result, large magnetic pressures and magnetic energy densities are present in tenuous layers, dominating over the gas pressure in the plasma equation of state and leading to numerical convergence errors. Models with initial masses above about 1.15~\msun\ are prevented from evolving through the final approach to the zero-age-main-sequence because of this problem. To mitigate the problem, a lower limit of $\sigma_g = 0.05$ is enforced.}

\begin{table}[t]
    \centering
    \caption{Equipartition magnetic field strength predictions at 10 Myr.}
    \begin{tabular*}{0.6\linewidth}{@{\extracolsep{\fill}}c c c}
        \hline\hline\noalign{\smallskip}
        Mass (\msun) & \teff (K) & $\langle Bf \rangle_{\rm eq,\,surf}$ (kG) \\
        \noalign{\smallskip}\hline\noalign{\smallskip}
        0.1  & 2920 & 2.58 \\
        0.2  & 3127 & 2.54 \\
        0.3  & 3260 & 2.56 \\
        0.4  & 3371 & 2.57 \\
        0.5  & 3463 & 2.56 \\
        0.6  & 3551 & 2.52 \\
        0.7  & 3640 & 2.47 \\
        0.8  & 3735 & 2.40 \\
        0.9  & 3836 & 2.29 \\
        1.0  & 3962 & 2.12 \\
        1.1  & 4099 & 1.94 \\
        1.2  & 4340 & 1.80 \\
        1.3  & 4617 & 1.69 \\
        1.4  & 4976 & 1.56 \\
        1.5  & 5565 & 1.29 \\
        1.6  & 6281 & 0.94 \\
        1.7  & 7004 & 0.77 \\
        \noalign{\smallskip}\hline
    \end{tabular*}
    \label{tab:equip_values}
\end{table}

% support with evidence from the literature? perhaps in discussion section?
The interior magnetic field strength profile requires a surface boundary condition: the average surface magnetic field strength $\langle{\rm B}f\rangle$. In previous studies, a range of values were explored and values that reproduce a given stellar fundamental property were eventually selected. 
Here, a different approach is used. Instead of adjusting models to find suitable magnetic field strengths, equipartition estimates of the maximum allowed average surface magnetic field strengths are adopted. The aim is to test the predictive abilities of magnetic stellar evolution models. Equipartition between the surface magnetic field pressure and the local gas pressure is assumed, meaning
\begin{equation}
    B_{\rm eq,\, surf} = \left(8\pi P_{\rm gas,\, surf}\right)^{1/2}.
\end{equation}
Surface gas pressures are defined using the same model atmosphere structures used to specify surface boundary conditions. The ``surface'' is defined as the optical photosphere where $\tau_{5000} = 1$. Values of the surface equipartition magnetic field strengths are listed in Table~\ref{tab:equip_values} for stars at an age of 10 Myr. Since surface gas pressures are dependent on \logg\ and \teff, equipartition magnetic field strengths evolve toward greater values for stars undergoing quasi-hydrostatic contraction. This assumption breaks down for stars with masses above 1.55 $M_{\odot}$, whose outer convection zones disappear within {\bf about} 10 Myr.

Magnetic field effects are switched on {\bf between the ages of 0.1 and 0.5 Myr years}, perturbing an initial zero-magnetic-field-strength model. The precise age at which the magnetic field is switched on has no impact on the results \citep{FC12b}, provided there is {\bf a} buffer between the perturbation age and ages under investigation (here, ages greater than 1 Myr). This buffer allows models to relax to a stable configuration after the perturbation. Typical peak magnetic field strengths in the models are of order 50 kG for models with $\langle Bf \rangle_{\rm eq,\,surf} \sim 2.5$~kG. 
{\bf Conversely, magnetic field effects are turned off during a model computation once the outer convective envelope disappears. This eventually occurs for stars with $M \gtrsim 1.15 M_{\odot}$.}

\begin{table}[t]
    \centering
    \caption{Adopted solar calibration properties.}
    \begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}l c c c}
        \hline\hline\noalign{\smallskip}
        Parameter & Adopted & Model & Ref \\
        \noalign{\smallskip}\hline\noalign{\smallskip}
        Age (Gyr)    & $4.567$ & $\cdots$ & 1 \\
        \msun\ (g)  & $1.9891\times 10^{33}$ & $\cdots$ & 2  \\
        \rsun\ (cm) & $6.9598\times 10^{10}$ & $\log(R/R_{\odot}) = -6\times 10^{-5}$ & 3, 1 \\
        \lsun\ (erg s$^{-1}$) & $3.8418\times 10^{33}$ & $\log(L/L_{\odot}) = -1\times 10^{-4}$ & 1 \\
        $R_{\rm bcz}/R_{\odot}$ & $0.713\pm0.001$ & $0.715$ & 4, 5 \\
        $(Z/X)_{\rm surf}$ & $0.0231$ & $0.0231$ & 6 \\
        $Y_{\odot,\, \rm surf}$ & $0.2485\pm0.0034$ & $0.2460$ & 7 \\
        \noalign{\smallskip}\hline
    \end{tabular*}
    \label{tab:solar}
    \tablebib{(1)~\citet{Bahcall2005};
    (2)~IAU 2009 \footnote{http://maia.usno.navy.mil/NSFA/NSFA\_cbe.html}
    (3)~\citet{Neckel1995}; (4)~\citet{Basu1997};~(5) \citet{Basu1998};
    (6)~\citet{GS98}; (7)~\citet{Basu2004}.}
\end{table}

Models for this study are computed using solar metallicity, [$m$/H] = 0. Values of the initial helium and heavy element mass fractions ($Y_0$ and $Z_0$, respectively) are taken from a solar-calibrated model. Similarly, a solar-calibrated mixing length parameter $\amlt$ is adopted for all models. In this way, modifications to convection are only introduced through the addition of magnetic perturbations. Solar-calibrated values are determined by finding a combination of $Y_0$, $Z_0$, and $\amlt$ that reproduce the solar radius, luminosity, and surface $\left(Z/X\right)$ at the solar age for a model of 1.0 \msun\ {\bf without a magnetic perturbation.} Adopted solar quantities are given in Table~\ref{tab:solar}. Final calibration parameters are $Y_0 = 0.2755$, $Z_0 = 0.01876$, and $\amlt = 1.884$.

\section{Properties of Upper Scorpius Members}
\label{sec:data}
Empirical data compared to stellar evolution models are taken from literature sources. Properties of stars with spectral type A, F, and G are drawn from \citet{Pecaut2012}. These data were kinematically selected to be highly probable members of Upper Scorpius \citep{deZeeuw1999}. In addition, data for K and M stars are drawn from the low-mass samples presented by \citet{Preibisch1999} and \citet{Preibisch2002}. Data span logarithmic effective temperatures from about 3.5 (3\,000~K) to 4.0 (10\,000~K). From these data, median empirical stellar loci were defined using a moving median approach.

A moving median of logarithmic luminosities as a function of $\log$(\teff) was performed by subdividing the temperature domain into 85 bins of equal width $\Delta\log$(\teff)$ = 0.025$~dex. {\bf This width ensured at least five stars were in each bin across the full temperature domain.} Bins with fewer than three stars were rejected. A median $\logLsun$\ value was estimated for each bin by finding the 50th percentile from the cumulative distribution of median values computed from a total of 10\,000 bootstrap samples. For each bootstrap sample, {\bf the $\log$(\teff) and $\logLsun$\ value for each data point} was randomly perturbed by drawing values from a normal distribution centered on the measured value with a standard deviation {\bf equal to} the quoted error. {\bf Thus}, points were allowed to scatter in and out of neighboring bins {\bf according} to their measured uncertainty. Points without uncertainties were assigned an {\bf optimistic uncertainty of 0.015~dex} in both $\logLsun$\ and $\log$(\teff). Different values for the number of bins, bin widths, and number of bootstrap samples were tested and found to not significantly impact the resulting median stellar locus. {\bf The resulting median locus and corresponding 99\% confidence interval for the median value are shown in Figure~\ref{fig:empirical}.}

\begin{figure}[t]
    \centering
    \includegraphics[width=0.85\linewidth]{USco_median_HRD.eps}
    \caption{Hertzsprung-Russell (HR) diagram for Upper Scorpius. A-type through M-type members of Upper Scorpius \citep{Preibisch1999,Preibisch2002,Pecaut2012} are shown as gray points. A median empirical locus for the population is shown as black solid line with the 99\% confidence interval for the median value shown as a light-gray shaded region. Errorbars shown for the majority of stars with \teff\ $< 5\,500$~K are prescribed as described in Sect.~\ref{sec:data}.}
    \label{fig:empirical}
\end{figure}

In addition to the photometric samples, properties of two eclipsing multiple-star systems recently observed by \emph{Kepler}/K2 are included. The first of these systems, UScoCTIO5, is a low-mass eclipsing binary (\porb$= 34$ days) with a $M_A = 0.329\pm0.002$\msun\ primary and a $M_B = 0.317\pm0.002$\msun\ secondary \citep{Kraus2015}. Radii were also measured and found to be $R_A = 0.834\pm0.006$\rsun\ and $R_B = 0.810\pm0.006$\rsun\ for the primary and secondary, respectively. Extracting reliable effective temperatures for components of eclipsing binary systems is notoriously difficult. However, the similarity between the two stars in UScoCTIO5 ($F_B/F_A \approx T_B/T_A \approx 1$) allows \citet{Kraus2015} to apply several different techniques for determining component effective temperatures, including directly determining the total system bolometric flux [\fbol$= (2.02\pm0.10)\times10^{-10}\textrm{ erg s}^{-1}$]. Assuming a distance of $145\pm15$~pc \citep{deZeeuw1999}, they estimate that $T_{{\rm eff}, A, B} = 3200\pm75$~K. {\bf Properties determined by \citet{David2015b} will be discussed in Sect.~\ref{sec:results}.}

The second system, HD 144548 (also, HIP 78977), was recently revealed to be a triply eclipsing hierarchical triple comprised of a solar-mass close binary system orbiting an F-type tertiary \citep{Alonso2015}. The inner binary has an orbital period \porb~$= 1.63$ days, compared to \porb~$= 33.9$ days for the orbit of the inner binary with the tertiary. Through a combination of photodynamical, radial velocity, and observed minus calculated ($O-C$) modeling, \citet{Alonso2015} were able to measure masses and radii for all three stars. They find $M_A = 1.44\pm0.04$\msun, $M_{Ba} = 0.984\pm0.007$\msun, and $M_{Bb} = 0.944\pm0.017$\msun\ for the tertiary and two close binary components, respectively. Radii for the stars were measured  to be $R_A = 2.41\pm0.03$\rsun, $R_{Ba} = 1.319\pm0.010$\rsun, and $R_{Bb} = 1.330\pm0.010$\rsun. While they were unable to estimate temperatures for individual stars, they were able to estimate luminosity ratios $L_{Bb}/L_{Ba} = 0.97\pm0.05$ and $L_{Ba + Bb}/L_{A} = 0.0593\pm0.0012$. In addition, they quote an apsidal motion rate $\dot{\omega} = 0.0235 \pm 0.002\textrm{ deg day}^{-1}$, which can potentially be used as an independent age estimator \citep{FD13}.

\citet{Pecaut2012} estimated the temperature and luminosity of HD 144548 as though it were a single star. For consistency, their values of $\log$(\teff) $ = 3.788\pm0.007$ and $\logLsun = 0.83\pm0.08$ are adopted as the temperature and luminosity of HD 144548 A. A 6\% correction was applied to the luminosity estimate to remove contribution from the close binary, as established by the luminosity ratio from photodynamical modeling \citep{Alonso2015}. Complications arising with this assignment of $\log$(\teff) and $\logLsun$ are discussed in Section~\ref{sec:radius}.

{\bf Fundamental properties for the eclipsing binary EPIC 203710387 \citep{David2015b, Lodieu2015} are not included in the main analysis because the component masses are uncertain at the 0.02 \msun\ level. Discussion with respect to model predictions in light the two fundamental property determinations will be presented in Sect.~\ref{sec:disc}.}

\begin{figure*}[t]
    \centering
    \includegraphics[width=0.48\linewidth]{HR_diagram_age_USco.eps} \quad
    \includegraphics[width=0.43\linewidth]{HR_diagram_split_USco.eps}
    \caption{{\bf (a) Hertzsprung-Russell (HR) diagram for Upper Scorpius. The median empirical locus calculated in Section~\ref{sec:data} is shown by the light-gray shaded region. Black points signify HR diagram positions of eclipsing binary members for which masses and radii are known. Over-plotted are predictions from Dartmouth stellar evolution isochrones. Standard (i.e., non-magnetic) isochrones at 5 Myr, 10 Myr, and 15 Myr are shown as short-dashed, dash-dotted, and long-dashed lines, respectively. A single magnetic isochrone at 10 Myr is plotted as a solid yellow line. All models are computed with a solar metallicity. (b) Zoom-in on the high-mass region of the HR diagram. (c) Zoom-in on the low-mass region.}}
    \label{fig:ages}
\end{figure*}

\section{Results}
\label{sec:results}
An HR diagram for Upper Scorpius is shown in Fig.~\ref{fig:ages}a. Empirical data is represented by the median stellar locus ({\bf light-gray shaded region}) and against three standard stellar evolution isochrones with ages of 5, 10, and 15 Myr ({\bf black dashed lines}). Focusing on {\bf temperature regimes corresponding} A-, F-, and G-type stars {\bf ($5\,750 <$ \teff/K $< 10\,000$)}, standard stellar isochrones predict ages between 5 and 15 Myr, with a median age estimate of approximately 10 Myr. %A median age of roughly 10 Myr is confirmed by comparison of the median stellar locus with stellar model isochrones.

Careful analysis of Fig.~\ref{fig:ages}b reveals that the empirical sequence closely matches models predictions at about 9 Myr where \teff~$> 8\,000$~K. Late A-type and early F-type stars with $7\,500 < $ \teff/K $ < 8\,000$ agree better with a 9 -- 10 Myr isochrone, as does the early-G-star sequence around \teff\ $\sim 5\,750$~K. In contrast, F-type stars with $6\,000 <$ \teff/K $< 7\,500$ appear marginally older and are best fit by an isochrone between 10 -- 15 Myr. These same features were noted by \citet{Pecaut2012} in their reanalysis of Upper Scorpius. Nevertheless, a 10 Myr isochrone lies almost entirely within the 99\% confidence interval for the median value across the entire range, suggesting an overall age of 10 Myr is reasonable. One can also see that HD 144548 A lies close to the 10 Myr isochrone suggesting that, at the very least, HD 144548 is approximately 10 Myr old.

However, the median age inferred from low-mass stars is about 4 -- 5 Myr. M-type stars with \teff\ $\lesssim 4\,000$~K suggest an age of 5 Myr is appropriate. The empirical stellar locus plotted in Figs.~\ref{fig:ages}a and \ref{fig:ages}c {\bf encompasses a 5 Myr standard stellar model isochrone in the M-star sequence down to \teff\ $\sim3\,100$ K. Disagreement at the lowest temperatures is likely due to a bias in the computation of the empirical stellar locus caused by a steepening of the HR diagram (see Figure~\ref{fig:empirical}).}

K-type stars with $4\,000 <$ \teff/K $< 5\,250$ suggest a younger age of around 4 Myr. Due to a lack of quoted uncertainties (see Sect.~\ref{sec:data}), it is difficult to rigorously assess the significance of the disagreement between the K- and M-star ages. The position of UScoCTIO5 in the HR diagram suggests an age that is slightly younger than 5 Myr, in line with the result of \citet{Kraus2015}. Both points for UScoCTIO5 lie just above the standard stellar model isochrone at 5 Myr, but within the empirical stellar locus \citep{Preibisch1999,Preibisch2002}. Its age is thus consistent with the 4 -- 5 Myr age inferred from K- and M-type stars.

A mass-radius diagram with the two eclipsing systems is shown in Fig.~\ref{fig:mrd}. Errorbars are plotted, but are generally smaller than the individual data points. It is apparent that ages inferred for the individual binary systems are not consistent with those from an HR diagram analysis. The age of HD 144548 from the Ba/Bb components is suggested to be around 5.5 Myr according to standard stellar evolution isochrones. While an age of 5.5 Myr agrees with quoted ages for the K- and M-type stars, it is in tension with the {\bf 10 Myr} age for component A from the HR diagram. Curiously, HD 144548 A lies significantly above the model mass-radius relation. Although concerning, further discussion of HD 144548 A's radius is deferred until Sect.~\ref{sec:radius}. The mass-radius diagram also suggests that UScoCTIO5 is 7.5 Myr old. This age is intermediate between the median age of the higher mass population and other low-mass stars, but it is important to note that this age is inconsistent with the HR diagram age of 5 Myr for UScoCTIO5.

\begin{figure}[t]
    \centering
    \includegraphics[width=0.85\linewidth]{USCO5_MR.eps}
    \caption{Mass-radius diagram for stars in eclipsing systems (black points). Model predictions for magnetic and non-magnetic isochrones are the same as in Fig.~\ref{fig:ages}.}
    \label{fig:mrd}
\end{figure}

Both eclipsing systems exhibit a disagreement between their HR and mass-radius diagram age. However, the disagreements do not immediately appear to be systematic. HD 144548 appears younger in the mass-radius plane than it does in the HR diagram by roughly a factor of two. In contrast, UScoCTIO5 appears older in the mass-radius plane than it does in the HR diagram---also by {\bf about} a factor of two. Slopes of the theoretical \teff-luminosity and mass-radius relationships must be altered in different directions to produce an overall agreement.

Introducing a magnetic perturbation as described in Sect.~\ref{sec:models}, a 10 Myr isochrone is computed and overlaid in the HR and mass-radius diagrams in Figs.~\ref{fig:ages} and \ref{fig:mrd} (thick yellow solid line). An age of 10 Myr was chosen as a starting point based on the approximate median age of the A-, F-, and G-type stars. Figures~\ref{fig:ages}a and \ref{fig:ages}c (HR diagrams) {\bf show that models}, at an age of 10 Myr with equipartition magnetic fields, are shifted toward cooler temperatures and slightly larger luminosities compared to a non-magnetic 10 Myr isochrone. Inhibition of convection by magnetic fields {\bf cools the stellar surface temperature thereby slowing} the contraction rate of young stars, which leads to stars having a larger radius and a larger luminosity at a given age. The combination of cooler surface temperatures and higher luminosities leads to a situation where a 10 Myr magnetic isochrone looks nearly identical to a non-magnetic isochrone at 5 Myr. In fact, the magnetic stellar model isochrone {\bf lies on top of} the 5 Myr standard model isochrone and {\bf within} the empirical stellar locus. Thus, a 10 Myr magnetic isochrone naturally explains the factor of two age difference observed between high- and low-mass stars, provided the high mass stars are not appreciably affected by non-standard physics (e.g., magnetic fields or rotation).

Furthermore, the mass-radius diagram in Figure~\ref{fig:mrd} shows a 10 Myr magnetic isochrone provides reasonable agreement with the two low-mass eclipsing binary systems. It is apparent that magnetic inhibition of convection leads to more significant radius ``inflation'' at higher masses as compared to lower masses. This effect leads to a change in slope of the model predicted mass-radius relationship that is consistent with observations. Although agreement is good, it is not perfect. {\bf Depending on whether mass and radius estimates are adopted from \citet{Kraus2015} or \citet{David2015b}, at 10 Myr the radii of the primary and secondary of UScoCTIO5 are either 3.2\% and 2.1\% \citeauthor{Kraus2015} larger than model predictions or 5.7\% and 6.6\% larger \citeauthor{David2015b}.} Similarly, models at 10 Myr predict radii that are 1.4\% too large and 1.5\% too small for the Ba and Bb components of HD 144548 at 10 Myr, respectively. An age between 9.0 and 9.5 Myr is preferred for UScoCTIO5, while 10 Myr is the best fit age for HD 144548 Ba/Bb.

This age spread could be indicative of an intrinsic age spread, but it is also possible {\bf to attribute this spread to errors in the mass and radius determinations of the eclipsing binary systems or to the fact} that the low-mass stars in UScoCTIO5 may require marginally stronger magnetic field strengths than are predicted from equipartition arguments. Nevertheless, it is encouraging that models provide this level of agreement with an age and surface magnetic field strengths determined \emph{a priori}.

\section{Discussion}
\label{sec:disc}
An age of {\bf 9 -- 10} Myr gives broad agreement across the HR and mass-radius diagrams. No stellar population in Upper Scorpius {\bf studied here robustly} supports the possibility that the median age of the association is 5 Myr old.\footnote{Ages for B-type members of Upper Scorpius suggest a range of ages from 2 -- 15 Myr depending on whether models account for rotation \citep{Pecaut2012}, providing some support for a younger age.} As noted above, F-type stars appear a few million years older than stars of other spectral types. This is concerning and should be explored further, especially considering that potential missing physics in the models would likely conspire to make F-stars older, not younger. Curiously, most F stars in Upper Scorpius are at a stage of evolution where radii increase and subsequently decrease over a period of roughly 4 Myr, eerily similar to the observed age discrepancy (see Fig.\ \ref{fig:bump} and text below for more details).

%While ages for the F stars remain in doubt, an overall age of 10 Myr provides good agreement with median age for B-type stars near the main-sequence turn-off \citep{Pecaut2012}.

\subsection{Magnetic Sequence Turn-off}
\label{sec:dynamo}
% F --> G type members are bright (V < 10 mag) and could be accessible to ZDI. Perhaps we'd see changes in magnetic strengths or topologies stretching from late-G to late-F.

Given that magnetic inhibition of convection is required to explain the anomalously young age of low-mass K- and M-type stars, it is worth considering at what mass or effective temperature the effects related to magnetism can be identified.
%begins to play a significant role in governing stellar structure. 
Looking at Fig.\ \ref{fig:ages}a, early- to mid-G stars follow a 10 Myr standard stellar model isochrone. Instead of continuing the trend, late-G stars appear to deviate toward a 5 Myr isochrone around \teff\ $\sim 5\,750$ K. %, evidenced by a flattening of the empirical locus. 
This may be an artifact of how the empirical locus was computed. However, it may indicate {\bf a transition} where the influence of magnetic inhibition of convection becomes identifiable.

{\bf Predictions from stellar evolution models suggest a transition should exist. Convective envelopes in stars with masses greater than $\sim1.2$ $M_{\odot}$ gradually recede from the deep interior toward the stellar surface in response to the ignition of the proton-proton ($p$-$p$) chain and continued release of gravitational potential energy as the star contracts. Eventually, the convective envelope in these stars will disappear entirely following ignition of the CN-cycle. As the convective envelope grows thinner, magnetic fields of a given strength will have a larger impact on the thermal structure of the super-adiabatic region. However, a restructuring of the super-adiabatic layers in a star with a thin convective envelope has less of an impact on the thermal structure of the deeper layers, which are assumed to be in radiative equilibrium, than for a star whose interior is largely convective. The thinning of the convective envelope is also accompanied by a drop in surface gas pressure as the star's radius inflates, lowering the permitted equipartition magnetic field strength. The combination of these two effects suggests that the transition from a magnetic sequence to a non-magnetic sequence should coincide with the disappearance of the convective envelope.} % confirm discussion of super-adiabatic layers

{\bf This observed transition is reproduced by the magnetic isochrone at 10 Myr, to some degree. For \teff\ $< 5\,000$~K, the magnetic isochrone closely matches predictions from a 5 Myr standard model isochrone, indicating magnetic inhibition of convection may be playing a role for stars at these temperatures. Compare this to the small segment of the magnetic isochrone above \teff\ $\sim 6\,000$~K, where it is beginning to converge with the 10 Myr standard model isochrone. Although the transition appears to occur over approximately the correct effective temperature domain, Fig.~\ref{fig:ages}a shows a sharp transition from a magnetic to a non-magnetic sequence between $5\,000 <$ \teff/K $< 6\,000$ whereas models predict a more gradual transition. In magnetic model space, this temperature domain corresponds to masses between $1.40$ and $1.56$ \msun. Inspection of Fig.~\ref{fig:mrd} reveals this mass range corresponds precisely to where the rapid thinning of the convection zone occurs, as expected.}

{\bf While this behavior is expected in the magnetic models, it does not need to necessarily correspond to the observed transition. General agreement between the behavior of the magnetic model isochrone and the observed stellar locus provides evidence that magnetic inhibition of convection is the mechanism responsible for the age discrepancy between the high- and low-mass stellar populations. However, it is clear that this statement only holds with respect to the plausible identification of a general mechanism with the models failing to reproduce the precise morphology at the magnetic sequence turn-off.}

If the age difference between warmer and cooler stars in Upper Scorpius is the result of magnetic inhibition of convection, the transition region between $5\,000 <$ \teff/K $< 6\,000$ provides an excellent laboratory for studying magnetic dynamo evolution as a function of convection zone depth and convective core size.

%If this is the case, it may help explain the significant gap between the end of the magnetic stellar evolution isochrone and the location of HD 144548 A in the HR diagram. The mass difference is only $0.16\pm0.04$ \msun, but the temperature difference is nearly 1600~K. By contrast, the expected difference is 500~K from standard models. It's interesting to note that the end of the magnetic 10 Myr isochrone is where stars are expected to begin a brief period of radius inflation prior to ignition of the CN cycle (see below). This is shown in Fig.~\ref{fig:bump}. Radius inflation occurs over a 3 -- 4 Myr timespan and may result in a gradual decrease of the magnetic field strength throughout the stellar envelope due to a roughly 30\% decrease in the envelope's angular velocity. The latter is due to an increase in the envelope's moment of inertia and may weaken magnetic dynamo processes. One might then expect a gradual reduction in the importance of magnetic inhibition of convection, creating a situation where stars in a narrow mass range span a relatively large range in effective temperature, resulting in the gap seen in Fig.~\ref{fig:ages}a.

Presence of magnetic and non-magnetic sequences has implications for the (sub)stellar initial mass function of Upper Scorpius \citep{Ardila2000, Preibisch2002}. Not only will magnetic inhibition of convection shift mass estimates for cool stars toward higher masses, but a gradually declining magnetic field in the regime of 1.4 \msun\ should result in a paucity of stars in the 4500~K -- 5500~K temperature range. Standard stellar initial mass functions \citep[e.g.,][]{Salpeter1955, Kroupa2002} predict a certain number of stars at a given mass relative to other masses. Having a range of magnetic field strengths in stars over such a narrow mass range (0.16 \msun) will spread stars of similar mass across a wider range in temperature. It's not clear from \citet{Preibisch2002} whether this paucity is observed. However, they do note a possible excess of low-mass stars compared to field initial mass functions \citep{Scalo1998, Kroupa2002} that appears consistent with predictions from magnetic models. The impact of magnetic inhibition of convection on the (sub)stellar initial mass function of Upper Scorpius is reserved for future work.

\begin{figure}
    \centering
    \includegraphics[width=0.85\linewidth]{radius_bump.eps}
    \caption{Brief period of radius inflation described in Sect.~\ref{sec:radius} for stars with masses of 1.60\msun\ (blue solid), 1.55\msun\ (green short-dashed), 1.50\msun\ (blue dash-dotted), 1.45\msun\ (yellow long-dashed). Mass tracks are computed without a magnetic field (i.e., non-magnetic configuration). Effective temperatures at these masses are characteristic of F stars at 10 Myr.}
    \label{fig:bump}
\end{figure}

\subsection{Radius of HD 144548 A}
\label{sec:radius}
Figure~\ref{fig:mrd} reveals that the radius of HD 144548 A is significantly larger than is expected from standard stellar evolutionary predictions. This poses a problem to claims of consistent ages across stellar populations in Upper Scorpius. However, it is not immediately clear that the observed radius error is a true astrophysical problem. For instance, adopting the effective temperature and luminosity estimate from \citet{Pecaut2012}, one derives a radius of $R = 2.2 \pm 0.2 R_{\odot}$. This is formally consistent with the radius derived by \citep{Alonso2015}, but the mean value is more consistent with estimates from stellar evolution isochrones around 10 Myr. Stellar models show an increase in stellar radius for stars with masses in the vicinity of $M = 1.5$\msun, illustrated {\bf as a function of stellar mass and age} in Fig.\ \ref{fig:bump}. As described in {\bf Sect.\ \ref{sec:dynamo}}, this radius increase is due to the convective envelope responding to increased energy generation from the $p$--$p$ chain and gravitational contraction of the core.

Assessing in detail whether the larger radius of HD 144548 A is real is beyond the scope of this work. However, if it requires revision, the result is unlikely to affect the inferred masses and radii derived for the close binary. Those largely depend on photodynamical modeling of the \emph{Kepler}/K2 lightcurve during events where Ba and Bb eclipse one another, not where the pair eclipse the tertiary component. Results in Sect.~\ref{sec:results} should be robust. Still, confirmation of the star's radius would be beneficial for future investigations to ensure there are no adverse effects on the radii of the B components.

{\bf HD 144548 appears to be in a fairly rapid phase of evolution given either the radius estimate from \cite{Pecaut2012} or \citet{Alonso2015}. This phase of evolution is sensitive to radiative opacities and nuclear reaction cross sections. Radiative opacities and the initial $p$-$p$ reaction cross section determine when radius inflation begins while CN-cycle reactions largely control when radius inflation ends.} It may be possible to use HD 144548 A to constrain these physics, notably nuclear reaction cross-sections for CN-cycle reactions, if the properties of HD 144548 A can be accurately constrained. CN-cycle reactions eventually generate enough energy to oppose core contraction and reverse the radius increase {\bf shown in Figs.\ \ref{fig:mrd} and \ref{fig:bump} \citep[also see][]{Stassun2014}}. Therefore, the maximum radius reached by a star during this phase of evolution is sensitive to the adopted CN-cycle cross sections.

Perhaps there are other physics missing for HD 144548 A that result in the radius discrepancy. Magnetic inhibition of convection {\bf does not appear to play a significant role in governing the radius at this phase of evolution. Convective envelopes contain very little mass ($M_{cz} < 10^{-3} M_{\odot}$) and models indicate that, during the brief period radius inflation, magnetic and non-magnetic mass tracks tend to converge (see Fig.~\ref{fig:mrd}).} However, the exact role that magnetic fields may {\bf play} in governing the pre-main-sequence contraction of young A-, F-, and G-type stars remains relatively unexplored. Attempts were made to include effects from magnetic fields on these stars, but numerical issues prevented models from evolving through CN-cycle ignition. A kink in the magnetic stellar model isochrone can be seen in Fig.\ \ref{fig:mrd} around 1.53 \msun\ highlighting where magnetic perturbations were switched off. This occurs before CN-cycle ignition is reached.

The numerical issue is related to the thinning of the convection zone and the method used to specify surface boundary conditions. 
Extrapolating from the impact of magnetic fields on lower-mass stars during the development of radiative cores, it is expected that there will be little change in the derived ages for higher mass stars, but confirmation is needed.

Effects of strong magnetic fields on convection in stellar cores are also neglected. Strong magnetic fields may be generated by the interaction of a fossil field and vigorous dynamo action \citep[e.g.,][]{Featherstone2009}. Core magnetic fields in present models are below 1 G, which is insufficient to influence core convection. The impact of significantly stronger magnetic fields on core convection and the subsequent effect on the properties of higher mass stars is reserved for a future investigation.

\subsection{On Starspots}
Starspots have been suggested as an alternative mechanism to slow the contraction of young low-mass stars \citep{Jackson2009, MM10, Jackson2014a, Somers2015b}. The primary difference between starspots and a more general magnetic inhibition of convection is that spots are assumed to be strong localizations covering some fraction of the stellar surface whereas magnetic inhibition of convection is assumed (for simplicity) to be globally pervasive. Starspots have qualitatively and quantitatively similar effects on stellar fundamental properties \citep[radius, \teff, luminosity;][]{Spruit1982a,Spruit1986,Somers2015b}, but differ in their impact on photometric properties. Colors of spotted stars are the result of a superposition of surface regions of different effective temperatures \citep{Spruit1986}, whereas a global inhibition of convection still assumes a single temperature optical photosphere \citep{Jackson2014a}.

Assuming starspots are analogous to sunspots, {\bf spots} can be understood to be local manifestations of convective inhibition near the optical photosphere \citep{Biermann1941,Deinzer1965}. The agreement between ages and stellar fundamental properties predicted by magnetic stellar evolution isochrones presented in this work with ages of higher mass stars in Upper Scorpius may indicate that a global inhibition of convection accurately captures relevant physics involved in halting contraction of low-mass stars. If starspots are also to be considered, one should find that young stars in Upper Scorpius have a high coverage fraction of spots ($f \sim 1$). In this limit, a global inhibition of convection and spotted models \citep[e.g.,][]{Somers2015b} should provide equivalent explanations for observed HR and mass-radius diagram discrepancies.

Photometric brightness modulation is observed in lightcurves of UScoCTIO5 and HD 144548, suggesting spots are present on their optical photospheres. However, it may be that a large fraction of the surface is covered by a non-zero magnetic field of near-equipartition value, accompanied by smaller regions of strong local magnetic flux concentration, such as observed on the young star \object{AD Leo} \citep{Shulyak2014}. The latter can lead to the appearance of cooler spots, although only covering a small fraction of the stellar surface. If spots have a small areal coverage, it is more likely that they will have a negligible impact on stellar structure as compared to global inhibition of convection from the more ubiquitous non-zero background magnetic field.

\begin{figure}[t]
    \includegraphics[width=0.85\linewidth]{Li7_USco_Beq.eps}
    \caption{Lithium abundance  as a function of effective temperature at 10 Myr. Predictions are shown from standard (light-blue dashed line) and magnetic (yellow solid line) stellar evolution isochrones. Models were computed using the setup described in Sect.~\ref{sec:models}. Lithium abundance is defined as $A(\,^7{\rm Li}) = \log_{10}(N_{\rm Li}/N_{\rm H}) + 12$.}
    \label{fig:lithium}
\end{figure}

\subsection{Lithium Depletion}
An approach to validating magnetic stellar models is the determination of lithium abundances {\bf \citep{MacDonald2015},} or observation of lithium equivalent widths. {\bf Magnetic inhibition of convection produces a cooler temperature structure in fully convective pre-main-sequence stars, thereby delaying lithium burning and preserving lithium to older ages \citep{Ventura1998, DAntona2000, MM10, FC13, Malo2014, MacDonald2015}. Starspots have a similar affect on lithium burning timescales \citep{Jackson2014b, Somers2015b}, which is perhaps not surprising given the deep connection between starspots and magnetic inhibition of convection \citep{Biermann1941, Deinzer1965}. Nevertheless, if magnetic fields are directly influencing the structure of young stars, they should leave an imprint on the observed lithium abundance distribution in young stellar populations.}

Figure \ref{fig:lithium} demonstrates the impact that magnetic inhibition of convection has on predicted lithium abundances for stars in Upper Scorpius at 10 Myr. The lithium depletion boundary at cooler temperatures remains largely unchanged. However, lithium depletion is not expected to occur in warmer stars with effective temperatures exceeding about 4500~K, assuming they possess a strong magnetic field. Therefore, K-stars are not expected to exhibit signatures of lithium depletion, in contrast to standard model predictions, which suggest stars with spectral types later than G are expected to show signatures of lithium depletion.

Stars with effective temperatures around 3550~K are predicted to exhibit the greatest amount of lithium depletion, as compared to a prediction of 3750~K for standard models. Some depletion is also expected to occur for stars with $3300 <$ \teff/K $< 3500$, which standard models suggest should not be observed. This prediction shows a strong similarity to trends in lithium equivalent widths for M-stars in Upper Scorpius \citep{Rizzuto2015}. The observed lithium equivalent width minimum is located between \teff\ = 3300 -- 3400~K \citep{Rizzuto2015}. While comparing lithium equivalent widths and predictions of lithium abundances is not completely correct, within the narrow range of temperatures surrounding the lithium abundance minimum, equivalent widths should accurately reflect trends in overall lithium abundances. Accounting for uncertainties in the SpT-\teff\ conversion, observed lithium equivalent widths support the validity of magnetic models and an inferred age of 10 Myr for the low-mass stars in Upper Scorpius.

\section{Summary \& Conclusions}
\label{sec:tellit}
Stellar evolution models and literature data are used to derive ages for stars with spectral types A through M in Upper Scorpius. I find a revised median HR diagram age of {\bf 9 -- 10} Myr for A-, F-, and G-type stars, with the exception of a population of F-type stars that appear 5 Myr older. These results agree with the revised $11\pm3$ Myr of \citet{Pecaut2012}. At the same time, I confirm the HR diagram age of 4 -- 5 Myr for K- and M-type stars using standard stellar evolution isochrones \citep{Preibisch2012, Slesnick2008, Herczeg2015}. However, isochrone ages of K- and M-type stars in eclipsing multiple-star systems observed by \emph{Kepler}/K2 \citep{Kraus2015, Alonso2015} disagree with the HR diagram analysis and suggest ages between 6 and 8 Myr for the same stars.

Application of stellar evolution models accounting for magnetic inhibition of convection \citep{FC12b, FC13} yields an age of {\bf 9 -- 10} Myr for K- and M-type stars in Upper Scorpius. Magnetic models provide a consistent explanation for the age discrepancy between high- and low-mass stars, as well as differences observed between HR diagram and mass-radius relationship ages. Magnetic models also reproduce the slope of the low-mass star mass-radius relationship. Critically, average surface magnetic field strengths are determined \emph{a priori} via equipartition arguments, highlighting the potential predictive power of magnetic stellar models of young stars. Furthermore, magnetic models around 10 Myr exhibit a lithium depletion pattern that is in agreement with inferences from observed lithium equivalent widths \citep{Rizzuto2015}.

Results from this work and \citet{Pecaut2012} strongly support adoption of an approximately 10 Myr median age for the Upper Scorpius subgroup of the Scorpius-Centaurus OB association. There is no population of stars in the Upper Scorpius that unambiguously support an age of 5 Myr. Still, additional eclipsing binary systems are needed to further populate the mass-radius relationship and solidify Upper Scorpius as a 10 Myr old stellar association.

Further work is needed to explore the impact of magnetic inhibition of convection on the (sub)stellar initial mass function for Upper Scorpius and other young stellar populations \citep[e.g.,][]{Ardila2000, Preibisch2002}. More work is also needed to explore whether magnetic inhibition of convection serves as a robust solution for noted age gradients as a function of \teff\ \citep{Naylor2009, Herczeg2015}. However, given the ease with which magnetic models appear to reconcile discrepancies in Upper Scorpius, the wider application of magnetic stellar models as a component of isochrone analyses is recommended. A small grid of magnetic stellar isochrones is under construction as is a web interface to request custom magnetic models.

\begin{acknowledgements}
The magnetic Dartmouth stellar evolution code was originally developed with support of National Science Foundation (NSF) grant AST-0908345. This work made use of NASA's Astrophysics Data System (ADS) and the VizieR catalogue access tool by CDS in Strasbourg, France. Figures in this manuscript were produced with Gnuplot 5 \citep{Gnuplot5.0}.
\end{acknowledgements}

\bibliography{papers}

\end{document}