### 2016-04-27 paper/ref/20160427_referee_response_01.txt
Resubmitted version

Dear Editor Bertout,

I have read and given considerable thought to the comments and questions posed by the reviewer. Much time has passed since the reviewer submitted the review, a fact that reflects the attention I gave to their comments and how much their comments influenced the further development of the magnetic models. Their emphasis that the intermediate mass star population was of importance was accurate, so I felt it wise to devote significant time and energy to ensuring that this region was well modeled. I believe I achieved this aim. Ultimately, I am very grateful for their excellent review.

The reviewers reading of the manuscript and comments concerning model convergence issues prompted a more extensive study regarding the effects of magnetic fields on the properties of the intermediate mass star population. During this study, I was able to uncover the source of the model's convergence issues, allowing a larger grid of magnetic stellar evolution models to be computed. This led to some interesting results. In particular, a significant amount of time was devoted to addressing disagreements between the morphology of the HR diagram in the region where surface convective envelopes are disappearing. This more extensive investigation led to significant revisions of the discussion section to address potential physical explanations for the observed disagreements and also physical explanation for notable attributes that are in agreement (convergence of magnetic model isochrones with standard model isochrones).

All changes made to the manuscript are marked with bold text and are summarized below.

Sincerely,

Gregory Feiden

----

Summary of key changes:

- The abstract was revised for clarity and two sentences were added to reflect results from the extension of models to the intermediate mass regime.

- A paragraph in the introduction mentioning lithium depletion was removed as it wasn't necessary for providing background to the problem.

- References to David et al. (2016) were added, including a couple sentence at the end of Section 3. This work appeared after initial submission, but contains data on the same eclipsing binary that was analyzed by Kraus et al. (2015).

- Changes to the stellar evolution made to mitigate convergence issues are discussed in Section 2.

- A log(g) column was added to Table 1.

- All figures were updated to include new models and to improve clarity.

- Section 4.2 now discusses the transition region in the HR diagram that arose from my extension of the magnetic models to higher masses.

- Section 5 includes more extensive discussion of B-type stars.

- Section 5.1 was added to discuss the physics underlying the transition region observed in the HR diagram.

- Section 5.2 was created from the end of the old Section 5.1.

- Section 5.3 is presented without discussion of model convergence issues and mentions carbon/nitrogen abundances, which also play a key role in governing the brief period of radius inflation.

- I've tried to clarify the discussion of starspots vs magnetic inhibition of convection.

- More references were added to the discussion of the lithium depletion boundary.

- The summary and conclusions section was expanded.

====

I would like thank the reviewer for a careful reading of the manuscript and their astute analysis of the convergence problems. Their comments and questions motivated a more detailed study of numerical convergence for the intermediate mass star population; a study that temporarily mitigated the original problem. This resulted in an interesting comparison of the intermediate mass star population in the region of the HR diagram where surface convection zones are expected to disappear. There may be very interesting physics occurring in this region that warrant further study. Section 5 has been extended to address interesting points that arose from this investigation.

>> 1. In the introduction and in the discussion of Li depletion, it is worthwhile to mention the recent paper on the same subject by MacDonald & Mullen (2015, MNRAS, 448, 2019).

Mention of lithium depletion in the introduction has been removed since it was not the primary focus of the paper. However, we did reference MacDonald & Mullan (2015) in the discussion of lithium depletion.

>> 2. **I appreciate that the following may be difficult to accomplish given the current state of the field, so I only ask the author to comment on the following, and discuss possible directions for future study.** The equipartition argument derives the surface gas pressure from atmosphere models that neglect magnetic fields. Whereas a self-consistent treatment would be more appropriate. What is required in order for the atmosphere and interior to be treated on equal footing w.r.t. magnetic field?

It's very true that it's difficult to address this problem given the current state of the field, although there is an effort being made to relieve this uncertainty.

To treat the atmosphere and interior on equal footing with respect to the magnetic field, one must either compute a new grid of atmosphere models with magnetic fields or one must incorporate detailed frequency-dependent radiative transfer in stellar evolution models. The former is intractable, as it would require a number of new atmosphere grids that are calculated with different magnetic field strengths.

It's my opinion that the second option is more viable, although difficult in practice. Incorporating detailed radiative transfer in stellar evolution calculations effectively requires model atmospheres to be computed simultaneously with stellar interior models. This is computationally expensive, but feasible. The difficulty is related to the best way of incorporating the magnetic field, such as how the magnetic field interacts with the plasma in the stellar atmosphere. Nevertheless, if this can be achieved, then it will be possible to calculate accurate surface boundary conditions on the fly including effects due to magnetic fields.

>> 3. In Section 3 we see (Figure 1) that the magnetic isochrone does not extend to such high masses as the standard models. We have to wait until Section 5 to learn why. It would be useful to the reader to have this information up front.

Agreed, information about model limitations should be presented earlier. They are now discussed in Section 2, although in a revised context since we were able to mitigate issues that prevented models from extending to higher masses (see point 5 below).

>> 4. I also note that the different panels of Figure 1 are not labelled although they are referred to as a,b,c in the caption and text.

Panels in Figure 1 are now labeled properly.

>> 5. The issue of what happens to higher mass stars than are modelled with magnetic fields in this paper is extremely important and I would encourage the author to discuss briefly what the issues with convergence are.  This will help the reader to understand the problem and may also help others who wish to work on magnetic models in the future as to what problems they may expect to encounter.  As I read it, it seems that the problem with surface magnetic fields occurs just as we would expect the surface convection zone to disappear above 1 Msol.

An astute analysis by the reviewer. I'll first note that the issue was largely mitigated. The problems and temporary solutions are discussed in Section 2.

The principle problem was, as the reviewer noted, the disappearance of the surface convection zone. Models presented by Feiden & Chaboyer (2013) linearly scaled the width of the Gaussian perturbation with the size of the surface convection zone. This prevented models of lower mass stars from having absurdly strong magnetic fields and allowed the magnetic field to vary more slowly throughout their deep surface convection zones. Model for solar-like stars, however, had a more localized (strongly peaked) magnetic field.

Issues with numerical convergence in the first version of the paper was a consequence of the Gaussian width going to zero, causing enormous magnetic field strengths near the tachocline, which was already situated close the stellar surface. The strong radial magnetic field strength gradient prevented the models from converging. By temporarily fixing the minimum value for the Gaussian width, I've avoided these numerical issues. The precise minimum value had little to no impact on the final results.

>> 6. To quote the estimated age of the system as 9-10 Myr, rather than 9.5 /-0.5 Myr, is perhaps more appropriate to the way the age is treated in the paper. We only see models with ages 5, 10, 15 Myr and no statistical analysis is done.

This is a fair point. The estimated median age is now quoted as 9 - 10 Myr.

>> 7. The stars above which no magnetic models are available are very close to the regime where they will arrive on the main sequence (as noted in the text to begin H-burning in the core), and since such models should have no surface convection zones on the main sequence (M > 1.3 Msol), we might benefit from a more detail about why the models fail to converge. This is certainly of interest to modellers who may try to conduct similar studies.

Agreed, but as mentioned in point 5, this problem has been temporarily relieved.

>> 8. The discussion of magnetic inhibited convection vs. star spots in curious to me because I would expect that they are each a manifestation of the same underlying phenomenon, i.e., magnetic fields in the outer layers of these stars. In this sense I find that the way the two are presented as distinct effects difficult. This also get back to point #2 above in which it would be very interesting to know whether both can be included in the same models.

Yes, this is an overly nuanced discussion. The reviewer is quite correct that starspots are the manifestation of magnetic inhibition of convection. They're discussed separately in Section 5 to highlight that theoretical formulations treat them as distinct phenomena. Magnetic inhibition of convection, in a stellar model sense, has only been treated as a global phenomenon (Mullan & MacDonald; Lydon & Sofia; Feiden & Chaboyer), where as starspots explicitly account for the level of surface coverage (Chabrier et al.; Jackson & Jeffries; Somers & Pinsonneault).

One might naively assume that magnetic inhibition of convection would produce more exaggerated results, since it's a global prescription. However, if magnetic inhibition of convection can reproduce the properties of stars with realistic magnetic field strengths, it suggests that the global description of the magnetic field is not a horrible assumption. Starspots models may then require high surface coverages (approaching 100%) in order to mimic the global prescription of models with magnetic inhibition of convection.

I've tried to make this more clear in Section 5.4.

>> 9. Can the author produce an age estimate from the decay of apsidal motion in HD 144548? If so, how does it compare with the other estimates?

In short, yes, but not reliably. An initial age estimate yielded an age around 15 - 20 Myr. While this is encouraging, it should be approached with caution.

The principle difficulty in deriving an age from the apsidal motion rate for HD 144548 is that the "secondary star" in the system is itself a binary. This poses problems when considering the age dating method proposed by Feiden & Dotter (2013). Namely, defining a suitable average interior structure constant (k2) for the inner binary that adequately describes the resulting gravitational potential, and defining a single "radius" for the inner binary.

>> 10. Finally, I note one reference that may be of interest on the subject of core magnetic fields from Kepler/K2, Fuller et al. 2015 (Science, 350, 423)

Yes, the paper by Fuller et al. on suppression of gravity waves in the cores of red giants is very interesting and quite relevant in terms of providing observational support for dynamo-generated magnetic fields in the cores of intermediate mass stars. The Fuller paper and two subsequent papers by the same group are now cited to motivate future inclusion of magnetic inhibition of convection in core convection descriptions.