### 2016-06-23 paper/ref/20160509_referee_response_02.txt
Edits and responses to report 2

> I have reviewed the latest draft of the paper and find it is essentially ready to be accepted. I would encourage the author to consider the following minor comments:

> - Section 5: I find the following two sentences difficult to understand. There is some confusion between real stars and models that needs to be clarified.

> "There are, however, still age discrepancies that must be addressed. As noted above, the median age for F-type stars appears to be a few million years older than stars of other spectral types. This discrepancy is perplexing and deserves further investigation, especially considering that potential missing physics in the models would likely conspire to make F-stars older, not younger."

I have attempted to emphasize a distinction between models and real stars by emphasizing that ages are inferred from models and are not a definite property of F-type stars.

> - Figure 4: It is much more common to discuss the mass of the convective envelope than of the radiative core.

This is true. I adopted the radiative core mass for consistency with Gregory et al. (2012), who used radiative core mass to explain variations in the geometry and complexity of large-scale magnetic fields among young stars. However, as noted by the reviewer, it is not an optimal selection, as convection zone structure is more pertinent for theoretical discussions of stellar structure and 3D magnetic dynamo simulations.

I've recast the discussion and Figure 4 in terms of convective envelope mass and refer to radiative core mass only for comparison with Gregory et al. (2012).